---
title: "Fitting Core STAN models"
author: "Chris Terry"
output: html_document
---


This script fits the core STAN models.

It also calculates the LOO likelihoods needed for elpd. 

It does not do any subsequent analysis - that is handled in `Main Text Analyses.rmd` or one of the SI documents. 

Assessment of sensitivity to priors is also conducted separately - SI 3.


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(rstan)
library(shinystan)
library(knitr)
library(scales)
library(loo)
library(cowplot)
# library(igraph)

source('Scripts/Stanfunctions.R') 

# 
# read_csv('Data/d_both.csv')%>%
#   select(- Temp,-Treatment,-MeanEmergence ,
#          -MeanGenTime,-ObsGrRate, - c(BIR:SUL),  -c(BIR_L:SUL_L), - Sp1_Dens) %>%
#   write_csv('Data/LongFormData.csv')


d_both<- read_csv('Data/LongFormData.csv')

species <- c('BIR','PAL', 'PAN', 'PSA', 'SIM', 'SUL')
Num_Draws <- 3000

dir.create('StanFits/')
dir.create('StanModels/')
dir.create('LOO/')
```

# Preparing data

NB '_Nt' values are inferred number of females values, so need to double up. 

```{r}
d_both <- mutate(d_both, ObsCount_Total =  ObsCount_Pairs*2)

STAN_data = list(N = nrow(d_both), 
                                      y = d_both$ObsCount_Total) 

```

# Writing and fitting STAN functions

## 1. Separate A and r models

```{r}
Write_Stan_NegB_2R_2A(d= d_both,
                      '2R_2A',
                      AlphaLowerBound = TRUE, AlphaSD= 1)
```

```{r}
fit_2R_2A <- stan( file = 'StanModels/BuiltModel_NegB_2R_2A.stan',
                          data = STAN_data, 
                          chains = 4, seed = 1,
                          cores = 2, iter = Num_Draws)

save(fit_2R_2A, file = 'StanFits/fit_2R_2A')
```

## 2. Single A, separate R

```{r}
Write_Stan_NegB_2R_1A(d= d_both,
                      '2R_1A',
                      AlphaLowerBound = TRUE, AlphaSD= 1)
```

```{r}
fit_2R_1A <- stan( file = 'StanModels/BuiltModel_NegB_2R_1A.stan',
                          data = STAN_data , 
                          chains = 4,  seed = 1,
                          cores = 2, iter = Num_Draws)


save(fit_2R_1A, file = 'StanFits/fit_2R_1A')
```

## 3. Single A and single r

```{r}
Write_Stan_NegB_1R_1A(d= d_both,
                      '1R_1A',
                      AlphaLowerBound = TRUE, AlphaSD= 1)
```

```{r}
fit_1R_1A <- stan( file = 'StanModels/BuiltModel_NegB_1R_1A.stan',
                             data = STAN_data , 
                             chains = 4,  seed = 1,
                             cores = 2, iter = Num_Draws)

save(fit_1R_1A, file = 'StanFits/fit_1R_1A')
```

# Cross-validation

```{r}
load('StanFits/fit_2R_2A')
load('StanFits/fit_2R_1A')
load('StanFits/fit_1R_1A')
```

```{r}
LOO_2R_2A<-loo(fit_2R_2A)
LOO_2R_1A<-loo(fit_2R_1A)
LOO_1R_1A<-loo(fit_1R_1A)
save(LOO_2R_2A, file='LOO/LOO_2R_2A')
save(LOO_2R_1A, file='LOO/LOO_2R_1A')
save(LOO_1R_1A, file='LOO/LOO_1R_1A')

```

