---
title: "Analyses Included in the Main Text"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(rstan)
library(shinystan)
library(knitr)
library(scales)
library(loo)
library(cowplot)
library(igraph)
library(gridExtra)
library(ggrepel)

source('Scripts/Other Functions.R')

species <- c('BIR','PAL', 'PAN', 'PSA', 'SIM', 'SUL')

cols = c('BIR' = 'skyblue', 
         'PAL' = 'darkblue',
         'PAN' = 'red',
         'PSA'= 'darkgreen',  
         'SIM' = 'purple', 
         'SUL' = 'gold'  )

# Loading model fits:

load('StanFits/fit_2R_2A')
load('StanFits/fit_2R_1A')
load('StanFits/fit_1R_1A')


draws_singleA<-extract(fit_2R_1A )

```

This document reproduces the analyses and figures in the main text of the paper. It is included for reproducibility, and is not an 'appendix' containing additional information as such. 

# Reproduction - susceptibility trade off

```{r warning = FALSE}

R_when_safe<- draws_singleA$r1
R_with_para <- draws_singleA$r2

CentralValue = data.frame(MedianR=  matrixStats::colMedians(R_when_safe),
                          MedianReductionPerc =matrixStats::colMedians(((R_when_safe)-(R_with_para))/(R_when_safe))*100,
                          Cols = cols,
                          Species = c("D. birchii",
                                      "D. pallidifrons",
                                      "D. pandora",
                                      "D. pseudoananassae",
                                      "D. simulans",
                                      "D. sulfurigaster"))

data.frame('Growth' =  as.vector(t(R_when_safe)),
           'Susceptibility'= as.vector(t(((R_when_safe)-(R_with_para))/(R_when_safe)))*100,
           'Species' = names(cols),
           'Cols' = cols)%>%
  arrange( Growth)%>%
  ggplot(aes(Growth,Susceptibility))+
  geom_point( aes( col = Cols ), alpha = 0.5, size = 0.5)+
 
  stat_ellipse(aes( col = Cols ), level = 0.6, size = 4, alpha = 0.5)+
  stat_ellipse(aes( group = Cols ), level = 0.6, size = 1, col = 'black')+
 geom_point(data = CentralValue,
             aes(MedianR, MedianReductionPerc, group = Cols ) , size =2)+
  geom_label_repel(data = CentralValue,
                   aes(MedianR, MedianReductionPerc,
                       label = Species ),
                   size =4, nudge_x = 0.015,
                   fontface = 'italic')+
  scale_color_identity()+
  scale_fill_identity()+
  xlab('Reproductive rate\nwithout parasitoid')+
  ylab("Susceptibility:\n% reduction in reproductive rate with parasitioid")+
   coord_cartesian(ylim = c(0,60),
                   xlim = c(0, 25))+
  theme_classic()

ggsave('Plots/Figure3.png', width = 6, height =4)
ggsave('Plots/Figure3.pdf', width = 6, height =4)


cor.test(CentralValue$MeanReductionPerc,
         CentralValue$MeanR,
         method = 'spearman')

cor.test(CentralValue$MeanReductionPerc,CentralValue$MeanR)

```


# Correlations between coefficents in fit
very strong positive correlations between r_i and a_ii. High estimated growth rates paired with high intra-specific competition terms. Also Correlations between two sets of R's, must be mediated by alpha's

```{r}
r1 <- draws_singleA$r1
r2 <- draws_singleA$r2 
As <- draws_singleA$A

r1_A_corr  <-  matrix(NA, 6,7)
for(i in 1:6){
  r1_A_corr[i,1:6]<-  cor(r1[,i], As[,((0:5)*6)+i])
  r1_A_corr[i,7]<-  cor(r1[,i], r2[,i])
  
}
colnames(r1_A_corr)<- c(paste0('a_i', 1:6 ), 'r2')
rownames(r1_A_corr)<- paste0('r_', 1:6)
kable(r1_A_corr, digits = 2)

r1_A_corr[c(1, 8, 15, 22, 29, 36)] 
mean(r1_A_corr[c(1, 8, 15, 22, 29, 36)] )

```

# Plotting Niche and Fitness Differences

Using Model 2 ('single A')

### Finding median values 

```{r}
r_safe <- matrixStats::colMedians(  draws_singleA$r1   )-1
r_para <- matrixStats::colMedians(  draws_singleA$r2   )-1
### Arranging 
A <-matrix( matrixStats::colMedians(  draws_singleA$A), 6,6)


Median_NicheOverlap = matrix(NA, 6,6)
Median_FitnessDifference_safe = matrix(NA, 6,6)
Median_FitnessDifference_para = matrix(NA, 6,6)

### Vectorised over draws (3rd dimension)
for(i in 1:6){
  for(j in 1:6){
    if(i ==j){ 
      Median_NicheOverlap[i,j]<- NA
      Median_FitnessDifference_para[i,j]<- NA
      Median_FitnessDifference_safe[i,j]<- NA
    }else{
      Median_NicheOverlap[i,j]<- sqrt(  (A[i,j]/A[j,j]) * (A[j,i]/A[i,i])   )
      Median_FitnessDifference_safe[i,j] <-   (r_safe[i]/r_safe[j])   *  sqrt(  (A[j,j]*A[j,i]) / (A[i,i]*A[i,j])) 
      Median_FitnessDifference_para[i,j] <-   (r_para[i]/r_para[j])   *  sqrt(  (A[j,j]*A[j,i]) / (A[i,i]*A[i,j])) 
    }
  }
}

Median_NicheDiff = 1- Median_NicheOverlap 

   
map_df(1:15, function(i,  Median_NicheDiff,  Median_FitnessDifference_safe,Median_FitnessDifference_para){

  ToSelect<-which(lower.tri(Median_FitnessDifference_safe), arr.ind = TRUE)
  
  df <- data.frame(Safe = Median_FitnessDifference_safe[ToSelect[i,1],ToSelect[i,2]] ,
                   Para = Median_FitnessDifference_para[ToSelect[i,1],ToSelect[i,2]] ,
                   NicheDiff =         Median_NicheDiff[ToSelect[i,1],ToSelect[i,2]])
  
  sp<- c('BIR','PAL', 'PAN', 'PSA', 'SIM', 'SUL')
  df$Combination <- paste0(sp[ToSelect[i,1]],'_',sp[ToSelect[i,2]])
  
  return(df)
},Median_NicheDiff,  Median_FitnessDifference_safe,Median_FitnessDifference_para)%>%
  gather( 'Treatment', FitnessDifference,  Safe:Para ) -> MedianValues

```

### Using full posterior

```{R}
HS_NiFi_draws<-CalcNicheFitDiffs_fromSTANfit(A_draws = draws_singleA$A,
                                             R_draws = draws_singleA$r1)
HP_NiFi_draws<-CalcNicheFitDiffs_fromSTANfit(A_draws = draws_singleA$A,
                                             R_draws = draws_singleA$r2)

AllValues_HS<- map_df(1:15, ExtractDiffs, HS_NiFi_draws   )
AllValues_HP<- map_df(1:15, ExtractDiffs, HP_NiFi_draws   )

bind_rows(mutate(AllValues_HS, Treatment = 'Safe'),
          mutate(AllValues_HP, Treatment = 'Para')) %>%
  mutate(Rho = 1- NicheDiff, 
         Priority         = 1/Rho < FitnessDiff &  FitnessDiff < Rho   &  NicheDiff <0,
         AboveBoundary_A1 = 1/Rho < FitnessDiff &  FitnessDiff > Rho   &  NicheDiff <0, 
         BelowBoundary_B1 = 1/Rho > FitnessDiff                        &  NicheDiff <0, 
         Coexistence      =   Rho < FitnessDiff &  FitnessDiff < 1/Rho &  NicheDiff >0,
         AboveBoundary_A2 =                        FitnessDiff > 1/Rho &  NicheDiff >0,
         BelowBoundary_B2 =   Rho > FitnessDiff                        &  NicheDiff >0, 
         Check = AboveBoundary_A1+AboveBoundary_A2+BelowBoundary_B1+BelowBoundary_B2+Coexistence+Priority
  ) -> LongFormPosterior


LongFormPosterior %>%
  group_by(Combination, Treatment)%>%
  summarise(MedianNicheDiff = median(NicheDiff),
            MedianFitnessDiff = median(FitnessDiff),
            MedianRho = median(Rho),
            MedianStable =  MedianRho<MedianFitnessDiff & MedianFitnessDiff<(1/(MedianRho))  & MedianNicheDiff>0,
            Above = mean(AboveBoundary_A1+AboveBoundary_A2)*100,
            Below = mean(BelowBoundary_B1+BelowBoundary_B2)*100, 
            Coexistence=mean(Coexistence)*100,
            Priority=mean(Priority)*100,
            Check = mean(Check))%>%
  separate(Combination , into = c('Species 1', 'Species 2'), remove = FALSE) -> PosteriorLocation

PosteriorLocation %>%
  left_join(MedianValues, by = c("Combination", "Treatment"))%>%
  arrange(Combination,desc(Treatment))%>%
  ungroup()%>%
  mutate( `Species 1` = recode(`Species 1`,
                               BIR = 'birchii',
                               PAL = 'pallidifrons',
                               PAN = 'pandora',
                               PSA = 'pseudoananassae',
                               SIM = 'simulans',
                               SUL = 'sulfurigaster')   ,
          `Species 2` = recode(`Species 2`,
                               BIR = 'birchii',
                               PAL = 'pallidifrons',
                               PAN = 'pandora',
                               PSA = 'pseudoananassae',
                               SIM = 'simulans',
                               SUL = 'sulfurigaster')   ) %>%
rename(`Species 1 Wins` = Above, 
         `Species 2 Wins` = Below ) %>%
  mutate(Difference = Coexistence-lag(Coexistence),
         MedianStable =(1-NicheDiff)<FitnessDifference   & FitnessDifference<(1/((1-NicheDiff)))  & NicheDiff>0 ) -> TidyTable


TidyTable%>% 
    select(-Combination,-Check, - (MedianNicheDiff :MedianRho ))%>%
  mutate(       Difference = ifelse(Treatment== 'Safe', ' ',round(Difference,1)  ),
         `Species 1`  = ifelse(Treatment== 'Para', ' ',`Species 1`  ),
         `Species 2`  = ifelse(Treatment== 'Para', ' ',  `Species 2`   ) )%>%
  kable(digits = 2) 

PosteriorLocation%>%
  select(Combination, Treatment , Coexistence)%>%
  spread(Treatment, Coexistence) %>% 
  ungroup()%>%
  mutate(Improved = Para>Safe)-> ImprovedOrNotData

ImprovedOrNotData%>%
  kable(digits = 1)

wilcox.test(ImprovedOrNotData$Para, ImprovedOrNotData$Safe, paired = TRUE) 

t.test(ImprovedOrNotData$Para, ImprovedOrNotData$Safe, paired = TRUE) 

PosteriorLocation%>%
  group_by(Treatment)%>%
  summarise(mean(Priority))

print('Overall proportion of negative niche differences:')
LongFormPosterior%>%
  summarise(mean(Rho>1))
```

# Coexistence Plane plots

```{R}
## Data pre-prep
ImprovedOrNotData%>%
  mutate(Shift = Safe-Para )%>%
  arrange(desc(Shift))-> OrderedCombs

bind_rows(mutate(AllValues_HS, Treatment = 'Safe'),
          mutate(AllValues_HP, Treatment = 'Para'))-> Both_TreatmentsValues

Both_TreatmentsValues$CombinationOrdered <- factor(Both_TreatmentsValues$Combination,
                                                   levels = OrderedCombs$Combination)

BoundaryValues<-data.frame(x = seq(0,0.99,l=100), 
                           y1 =   1/ (1-seq(0,0.99,l=100)),
                           y2 = (1-seq(0,0.99,l=100)))      


BoundaryValues_P<-data.frame(x = seq(0,-0.5,l=100), 
                             y1 =   1/ (1-seq(0,-0.5,l=100)),
                             y2 = (1-seq(0,-0.5,l=100)))      

```

## Explanatory diagram

```{r}
OneLine<- data.frame(x= c(rev(BoundaryValues$x),BoundaryValues$x),
                     y= c(rev(BoundaryValues$y2),BoundaryValues$y1 ))

Labels <- data.frame(Text = c('Priority\neffect',
                              'Species i wins',
                              'Species j wins',
                              'Coexistence'),
                     x = c(-0.3, 0.1, 0.1, 0.5),
                     y = c(1.1, 3, 0.3, 1))

BoundaryValues%>%
  ggplot()+
  geom_vline(xintercept = 0, linetype = 'dashed', col = 'grey')+
  geom_ribbon(data = BoundaryValues,
              aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
  geom_ribbon(data = BoundaryValues_P,
              aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
  scale_y_log10(name = expression(paste("Fitness difference: ",kappa[i]/kappa[j])) )+
  xlab(expression(paste("Niche Difference: ", 1-rho)))+
  coord_fixed(xlim =c(-0.5, 1), ylim= c(0.25, 4))+
  geom_path(data = OneLine,
            aes(x, y), col = 'red', size = 1.5)+
  geom_line(data = BoundaryValues,
            aes(x, y2), col = 'red', size = 1.5)+
  geom_text(data=Labels, aes(x,y, label =Text), size = 4)+
  scale_fill_viridis_c(guide = FALSE )+
  theme(strip.background = element_rect(fill = 'white'),
        strip.placement = 'left',
        strip.text = element_text(hjust = 0.1, vjust = 1),
        panel.spacing = unit(2, 'lines'))+
  theme_classic()-> Fig2_Explainer

Fig2_Explainer

```

## Demonstration Species pair

```{r}
SpeciesToPlot <- 'SUL_BIR'

Both_TreatmentsValues %>%
  filter( Combination  ==SpeciesToPlot ,
          Treatment == 'Para') %>%
  ggplot(aes(NicheDiff,FitnessDiff ))+
  geom_vline(xintercept = 0, linetype = 'dashed')+
  geom_ribbon(data = BoundaryValues,
              aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
  geom_ribbon(data = BoundaryValues_P,
              aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
  stat_density_2d(aes(fill = ..level..), geom = "polygon", bins = 8)+
  scale_y_log10(name = expression(paste("Fitness difference: ",kappa[SUL]/kappa[BIR])) )+
  xlab(expression(paste("Niche Difference: ", 1-rho)))+
  coord_fixed(xlim =c(-0.25, 1), ylim= c(0.8, 5))+
  geom_path(data = OneLine,
            aes(x, y), col = 'red', size = 1.5)+
  geom_line(data = BoundaryValues,
            aes(x, y2), col = 'red', size = 1.5)+
  scale_fill_gradient(low = 'bisque', high= 'darkred', guide = FALSE,  )+
  theme_classic()-> focal_Para

Both_TreatmentsValues %>%
  filter( Combination  ==SpeciesToPlot ,
          Treatment == 'Safe') %>%
  ggplot(aes(NicheDiff,FitnessDiff ))+
  geom_vline(xintercept = 0, linetype = 'dashed')+
  geom_ribbon(data = BoundaryValues,
              aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
  geom_ribbon(data = BoundaryValues_P,
              aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
  stat_density_2d(aes(fill = ..level..), geom = "polygon", bins = 8)+
  scale_y_log10(name = expression(paste("Fitness difference: ",kappa[SUL]/kappa[BIR])) )+
  xlab(expression(paste("Niche Difference: ", 1-rho)))+
  coord_fixed(xlim =c(-0.25, 1), ylim= c(0.8, 5))+
  geom_path(data = OneLine,
            aes(x, y), col = 'red', size = 1.5)+
  geom_line(data = BoundaryValues,
            aes(x, y2), col = 'red', size = 1.5)+
  scale_fill_gradient(low = 'slategray1', high= 'blue', guide = FALSE,  )+
   theme_classic()-> focal_Safe



```


## Contour Plots
```{r fig.height=10, fig.width=6, warning=FALSE}
TidyName<-  function(name){
  parts<- str_split_fixed(name, '_', 2)  
  nicename<- paste0('italic(i):~', parts[,1], '~italic(j):~',  parts[,2])
  return(nicename)
}

sorter<-   mutate(OrderedCombs, Combination=TidyName(Combination) )
sorter$numerals <- str_replace_all(formatOL(sorter$Combination, type = 'roman'), pattern = ' ', replacement = '~')
```


```{R}

Both_TreatmentsValues %>%
  mutate(Combination=TidyName(Combination) )%>%
  left_join(sorter, by = "Combination") %>%
  filter(NicheDiff>-1)%>%
  arrange(NicheDiff )%>%  # to blend dots better
  ggplot(aes(NicheDiff,FitnessDiff ))+
  geom_ribbon(data = BoundaryValues,
              aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
  geom_ribbon(data = BoundaryValues_P,
              aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
  geom_point(aes(col = Treatment), alpha = 0.05, size = 0.5)+
  geom_point(aes(fill = Treatment, NicheDiff, FitnessDifference), shape = 23, col = 'yellow',
             data = mutate(MedianValues,Combination=TidyName(Combination) ))+
  scale_y_log10(labels = function(x) sprintf("%g", x))+
  scale_colour_manual(values = c('red', 'blue'),
                      labels = c('With Parasitoid', 'No Parasitoid'),
                      name = 'Treatment')+
  guides(fill = FALSE, col = FALSE )+
  facet_wrap(~Combination, ncol = 3, labeller = label_parsed)+
  coord_cartesian(xlim = c(-0.25, 1), ylim = c(0.01, 100))+
  ylab(expression(paste("Fitness difference: ",kappa[i]/kappa[j])))+
  xlab(expression(paste("Niche Difference: ", 1-rho)))+
  theme_classic()+
  theme(strip.background = element_rect(fill = 'white', color = 'white', ),
        strip.text = element_text(hjust = 0.1, vjust = 1),
        panel.background = element_rect(colour = 'black'),
        panel.spacing = unit(2, 'lines'),
        axis.line = element_blank(),
        legend.position = "bottom")-> ContourPlot

ContourPlot

```

```{r fig.height=10, fig.width=10}
### Combining plots to make figure 3

CombinedPlot<- grid.arrange(Fig2_Explainer + ggtitle( 'a) Guide to coexistence plane'),
                            focal_Safe + ggtitle( expression(paste("     b) Example:",
                                                                   italic("D. sulfurigaster"),
                                                                   ' & ',
                                                                   italic("D. birchii"))),
                                                  
                                                  subtitle =  'i) Without Parasitoids') ,
                            focal_Para + ggtitle( '',subtitle ='ii) With Parasitoids'),
                            ContourPlot +   ggtitle('c) All Species Combinations'), 
                            layout_matrix =  rbind(c(1,1,4,4,4),
                                                   c(1,1,4,4,4),
                                                   c(2,2,4,4,4),
                                                   c(2,2,4,4,4),
                                                   c(3,3,4,4,4),
                                                   c(3,3,4,4,4)))
ggsave('Plots/CoexistencePlots.png',CombinedPlot, width = 10, height = 10)
ggsave('Plots/CoexistencePlots.pdf',CombinedPlot, width = 10, height = 10)

```


### Usefullness of fecundity - susceptibility tradeoffs

Assessment of reproductive rate – susceptibility relationships as a heuristic for the impact of parasitism on pairwise coexistence. 7/15 were in the ‘correct’ direction, strongly suggesting that such tradeoffs are not useful shortcuts ion practice, once the impact of competitive coefficients are taken into account. 

```{r}
Susceptibility = ((r_safe)-(r_para))/(r_safe)*100
# Might naively expect parasitism to aid coexistence if the more fecund species is also more susceptible... 

ExpectParaToHelp = matrix(NA, 6,6)

### Vectorised over draws (3rd dimension)
for(i in 1:6){
  for(j in 1:6){
    if(i ==j){ 
      ExpectParaToHelp[i,j]<- NA
    }else{
      if(r_safe[i]>r_safe[j]){   ExpectParaToHelp[i,j]<-  Susceptibility[i]>Susceptibility[j] }
      if(r_safe[i]<r_safe[j]){   ExpectParaToHelp[i,j]<-  Susceptibility[i]<Susceptibility[j]  }
    }
  }
}

colnames( ExpectParaToHelp) <- species
rownames( ExpectParaToHelp) <- species

ExpectParaToHelp %>%
  as.data.frame()%>%
  rownames_to_column('sp1') %>%
  gather( 'sp2', 'help', BIR:SUL) %>%
  mutate( Combination = paste(sp1, sp2, sep = '_')) %>%
  right_join(TidyTable, by = "Combination") %>%
  mutate( FitnessDifferenceReduced = ((FitnessDifference-1)- lag( FitnessDifference-1))<0) %>%
  filter( Treatment == 'Para') %>% 
  select(`Species 1`,`Species 2` , ParaExpectedToHelp = help,  FitnessDifferenceReduced) %>%
  mutate( CorrectDirection = ParaExpectedToHelp == FitnessDifferenceReduced) %>%
  kable

```



