---
title: "Analyses Included in the Main Text"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(rstan)
library(shinystan)
library(knitr)
library(scales)
library(loo)
library(cowplot)
library(igraph)
library(ggrepel)

source('Scripts/Other Functions.R')

species <- c('BIR','PAL', 'PAN', 'PSA', 'SIM', 'SUL')

cols = c('BIR' = 'skyblue', 
         'PAL' = 'darkblue',
         'PAN' = 'red',
         'PSA'= 'darkgreen',  
         'SIM' = 'purple', 
         'SUL' = 'gold'  )

# Loading model fits:

load('StanFits/fit_both_singleA')
load('StanFits/fit_both_doubleA')
load('StanFits/fit_both_Comb_RandA')


draws_singleA<-extract(fit_both_singleA )

```

This document reproduces the analyses and figures in the main text of the paper. It is included for reproducibility, and is not an 'appendix' containing additional information as such. 

# Reproduction - susceptibility trade off

```{r warning = FALSE}

R_when_safe<- draws_singleA$r1
R_with_para <- draws_singleA$r2

CentralValue = data.frame(MeanR=colMeans(R_when_safe),
                          MeanReductionPerc =colMeans(((R_when_safe-1)-(R_with_para-1))/(R_when_safe-1))*100,
                          Cols = cols,
                          Species = c("D. birchii",
                                      "D. pallidifrons",
                                      "D. pandora",
                                      "D. pseudoananassae",
                                      "D. simulans",
                                      "D. sulfurigaster"))

data.frame('Growth' =  as.vector(t(R_when_safe)),
           'Defense'= as.vector(t(((R_when_safe-1)-(R_with_para-1))/(R_when_safe-1)))*100,
           'Species' = names(cols),
           'Cols' = cols)%>%
  ggplot(aes(Growth,Defense))+
  stat_ellipse(aes( col = Cols ))+
  geom_point( aes( col = Cols ), alpha = 0.1, size = 0.1)+
  geom_point(data = CentralValue,
             aes(MeanR, MeanReductionPerc, fill = Cols ) , size =2, pch = 21)+
  geom_label_repel(data = CentralValue,
                   aes(MeanR, MeanReductionPerc,
                       label = Species ),
                   size =4, nudge_x = 0.015,
                   fontface = 'italic')+
  scale_color_identity()+
  scale_fill_identity()+
  xlab('Fecundity\nGrowth rate (r) without parasitoid')+
  ylab('Susceptibility\n% reduction in rate of\nincrease (r-1) with parasitioid ')+
  coord_cartesian(ylim = c(0,60),
                  xlim = c(1.05, 1.35))


ggsave('Plots/Figure2.png', width = 6, height =4)
ggsave('Plots/Figure2.pdf', width = 6, height =4)


cor.test(CentralValue$MeanReductionPerc,
         CentralValue$MeanR,
         method = 'spearman')

cor.test(CentralValue$MeanReductionPerc,CentralValue$MeanR)

```


# Correlations between coefficents in fit
very strong positive correlations between r_i and a_ii. High estimated growth rates paired with high intra-specific competition terms. Also Correlations between two sets of R's, must be mediated by alpha's

```{r}
draws_singleA<-extract(fit_both_singleA )

r1 <-draws_singleA$r1
r2 <- draws_singleA$r2 
As <-draws_singleA$A

r1_A_corr  <-  matrix(NA, 6,7)
for(i in 1:6){
  r1_A_corr[i,1:6]<-  cor(r1[,i], As[,((0:5)*6)+i])
  r1_A_corr[i,7]<-  cor(r1[,i], r2[,i])
  
}
colnames(r1_A_corr)<- c(paste0('a_i', 1:6 ), 'r2')
rownames(r1_A_corr)<- paste0('r_', 1:6)
kable(r1_A_corr, digits = 2)

r1_A_corr[c(1, 8, 15, 22, 29, 36)] 
mean(r1_A_corr[c(1, 8, 15, 22, 29, 36)] )

```


```{r}
### Structure in alpha matrix
mean_alpha_mat<-matrix(colMeans(As), 6,6)
colMeans(mean_alpha_mat)*1000
rowMeans(mean_alpha_mat)*1000
```

# Plotting Niche and Fitness Differences

Using Model 2 ('single A')



```{R}
HS_NiFi_draws<-CalcNicheFitDiffs_fromSTANfit(A_draws = draws_singleA$A,
                                             R_draws = draws_singleA$r1)
HP_NiFi_draws<-CalcNicheFitDiffs_fromSTANfit(A_draws = draws_singleA$A,
                                             R_draws = draws_singleA$r2)

AllValues_HS<- map_df(1:15, ExtractDiffs, HS_NiFi_draws   )
AllValues_HP<- map_df(1:15, ExtractDiffs, HP_NiFi_draws   )

bind_rows(mutate(AllValues_HS, Treatment = 'Safe'),
          mutate(AllValues_HP, Treatment = 'Para')) %>%
  mutate(Rho = 1- NicheDiff, 
         Priority         = 1/Rho < FitnessDiff &  FitnessDiff < Rho   &  NicheDiff <0,
         AboveBoundary_A1 = 1/Rho < FitnessDiff &  FitnessDiff > Rho   &  NicheDiff <0, 
         BelowBoundary_B1 = 1/Rho > FitnessDiff                        &  NicheDiff <0, 
         Coexistence      =   Rho < FitnessDiff &  FitnessDiff < 1/Rho &  NicheDiff >0,
         AboveBoundary_A2 =                        FitnessDiff > 1/Rho &  NicheDiff >0,
         BelowBoundary_B2 =   Rho > FitnessDiff                        &  NicheDiff >0, 
         Check = AboveBoundary_A1+AboveBoundary_A2+BelowBoundary_B1+BelowBoundary_B2+Coexistence+Priority
  ) -> LongFormPosterior



LongFormPosterior %>%
  group_by(Combination, Treatment)%>%
  summarise(MedianNicheDiff = median(NicheDiff),
            MedianFitnessDiff = median(FitnessDiff),
            MedianRho = median(Rho),
            MedianStable =  MedianRho<MedianFitnessDiff & MedianFitnessDiff<(1/(MedianRho))  & MedianNicheDiff>0,
            Above = mean(AboveBoundary_A1+AboveBoundary_A2)*100,
            Below = mean(BelowBoundary_B1+BelowBoundary_B2)*100, 
            Coexistence=mean(Coexistence)*100,
            Priority=mean(Priority)*100,
            Check = mean(Check))%>%
  separate(Combination , into = c('Species 1', 'Species 2'), remove = FALSE) -> PosteriorLocation


PosteriorLocation%>%
  arrange(Combination,desc(Treatment))%>%
  ungroup()%>%
  mutate( `Species 1` = recode(`Species 1`,
                               BIR = 'birchii',
                               PAL = 'pallidifrons',
                               PAN = 'pandora',
                               PSA = 'pseudoananassae',
                               SIM = 'simulans',
                               SUL = 'sulfurigaster')   ,
          `Species 2` = recode(`Species 2`,
                               BIR = 'birchii',
                               PAL = 'pallidifrons',
                               PAN = 'pandora',
                               PSA = 'pseudoananassae',
                               SIM = 'simulans',
                               SUL = 'sulfurigaster')   ) %>%
  select(-Combination,-Check, - MedianRho,
         `Species 1 Wins` = Above, 
         `Species 2 Wins` = Below)%>%
  mutate(Difference = Coexistence-lag(Coexistence))%>%
  mutate(Difference = ifelse(Treatment== 'Safe', ' ',round(Difference,1)  ),
         `Species 1`  = ifelse(Treatment== 'Para', ' ',`Species 1`  ),
         `Species 2`  = ifelse(Treatment== 'Para', ' ',  `Species 2`   )) %>% 
  kable(digits = 2) 

PosteriorLocation%>%
  select(Combination, Treatment , Coexistence)%>%
  spread(Treatment, Coexistence) %>% 
  ungroup()%>%
  mutate(Improved = Para>Safe)-> ImprovedOrNotData

ImprovedOrNotData%>%
  kable(digits = 1)

wilcox.test(ImprovedOrNotData$Para, ImprovedOrNotData$Safe, paired = TRUE) 

t.test(ImprovedOrNotData$Para, ImprovedOrNotData$Safe, paired = TRUE) 

PosteriorLocation%>%
  group_by(Treatment)%>%
  summarise(mean(Priority))

print('Overall proportion of negative niche differences:')
LongFormPosterior%>%
  summarise(mean(Rho>1))
```

# Figure 3: Coexistence Plante plots

```{R}
## Data pre-prep
ImprovedOrNotData%>%
  mutate(Shift = Safe-Para )%>%
  arrange(desc(Shift))-> OrderedCombs

bind_rows(mutate(AllValues_HS, Treatment = 'Safe'),
          mutate(AllValues_HP, Treatment = 'Para'))-> Both_TreatmentsValues

Both_TreatmentsValues$CombinationOrdered <- factor(Both_TreatmentsValues$Combination,
                                                   levels = OrderedCombs$Combination)

BoundaryValues<-data.frame(x = seq(0,0.99,l=100), 
                           y1 =   1/ (1-seq(0,0.99,l=100)),
                           y2 = (1-seq(0,0.99,l=100)))      


BoundaryValues_P<-data.frame(x = seq(0,-0.5,l=100), 
                             y1 =   1/ (1-seq(0,-0.5,l=100)),
                             y2 = (1-seq(0,-0.5,l=100)))      

```

## Explanatory diagram

```{r}
OneLine<- data.frame(x= c(rev(BoundaryValues$x),BoundaryValues$x),
                     y= c(rev(BoundaryValues$y2),BoundaryValues$y1 ))



Labels <- data.frame(Text = c('Priority\neffect',
                              'Species i wins',
                              'Species j wins',
                              'Coexistence'),
                     x = c(-0.3, 0.1, 0.1, 0.5),
                     y = c(1.1, 3, 0.3, 1))

BoundaryValues%>%
  ggplot()+
  geom_vline(xintercept = 0, linetype = 'dashed', col = 'grey')+
  geom_ribbon(data = BoundaryValues,
              aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
  geom_ribbon(data = BoundaryValues_P,
              aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
  scale_y_log10(name = expression(paste("Fitness difference: ",kappa[i]/kappa[j])) )+
  xlab(expression(paste("Niche Difference: ", 1-rho)))+
  coord_fixed(xlim =c(-0.5, 1), ylim= c(0.25, 4))+
  geom_path(data = OneLine,
            aes(x, y), col = 'red', size = 1.5)+
  geom_line(data = BoundaryValues,
            aes(x, y2), col = 'red', size = 1.5)+
  geom_text(data=Labels, aes(x,y, label =Text), size = 4)+
  scale_fill_viridis_c(guide = FALSE )+
  theme(strip.background = element_rect(fill = 'white'),
        strip.placement = 'left',
        strip.text = element_text(hjust = 0.1, vjust = 1),
        panel.spacing = unit(2, 'lines'))-> Fig2_Explainer

Fig2_Explainer

```

## Demonstration Species pair

```{r}
SpeciesToPlot <- 'SUL_SIM'

Both_TreatmentsValues %>%
  filter( Combination  ==SpeciesToPlot )%>%
  mutate(Treatment = recode(Treatment, 
                            Para = 'ii) With Parasitoids',
                            Safe = 'i) Without Parasitoids') )%>%
  ggplot(aes(NicheDiff,FitnessDiff ))+
  geom_vline(xintercept = 0, linetype = 'dashed')+
  geom_ribbon(data = BoundaryValues,
              aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
  geom_ribbon(data = BoundaryValues_P,
              aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
  stat_density_2d(aes(fill = ..level..), geom = "polygon", bins = 8)+
  scale_y_log10(name = expression(paste("Fitness difference: ",kappa[i]/kappa[j])) )+
  xlab(expression(paste("Niche Difference: ", 1-rho)))+
  coord_fixed(xlim =c(-0.25, 1), ylim= c(0.2, 3))+
  geom_path(data = OneLine,
            aes(x, y), col = 'red', size = 1.5)+
  geom_line(data = BoundaryValues,
            aes(x, y2), col = 'red', size = 1.5)+
  scale_fill_viridis_c(guide = FALSE )+
  geom_point(data = filter(PosteriorLocation,  Combination  ==SpeciesToPlot )%>%
               mutate(Treatment = recode(Treatment, 
                                         Para = 'ii) With Parasitoids',
                                         Safe = 'i) Without Parasitoids') ),
             aes(MedianNicheDiff,MedianFitnessDiff ), col = 'black', shape  =4)+
  facet_wrap(~Treatment, ncol=1)+
  theme(strip.background = element_rect(fill = 'white'),
        strip.placement = 'left',
        strip.text = element_text(hjust = 0.1, vjust = 1),
        panel.spacing = unit(2, 'lines'))-> SUL_SIMContourplot

SUL_SIMContourplot+ 
  ggtitle('b) Example Pair:\nD. sulfurigaster and D. simulans')+
  theme(title = element_text(hjust = 0))

```

## Using different colour schemes


```{r}

Both_TreatmentsValues %>%
  filter( Combination  ==SpeciesToPlot ,
          Treatment == 'Para') %>%
  mutate(Treatment = recode(Treatment, 
                            Para = 'ii) With Parasitoids',
                            Safe = 'i) Without Parasitoids') )%>%
  ggplot(aes(NicheDiff,FitnessDiff ))+
  geom_vline(xintercept = 0, linetype = 'dashed')+
  geom_ribbon(data = BoundaryValues,
              aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
  geom_ribbon(data = BoundaryValues_P,
              aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
  stat_density_2d(aes(fill = ..level..), geom = "polygon", bins = 8)+
  scale_y_log10(name = expression(paste("Fitness difference: ",kappa[i]/kappa[j])) )+
  xlab(expression(paste("Niche Difference: ", 1-rho)))+
  coord_fixed(xlim =c(-0.25, 1), ylim= c(0.2, 3))+
  geom_path(data = OneLine,
            aes(x, y), col = 'red', size = 1.5)+
  geom_line(data = BoundaryValues,
            aes(x, y2), col = 'red', size = 1.5)+
  scale_fill_gradient(low = 'darkred', high= 'white', guide = FALSE,  )+
  facet_wrap(~Treatment, ncol=1)+
  theme(strip.background = element_rect(fill = 'white'),
        strip.placement = 'left',
        strip.text = element_text(hjust = 0.1, vjust = 1),
        panel.spacing = unit(2, 'lines'))-> SULSIM_Para

Both_TreatmentsValues %>%
  filter( Combination  ==SpeciesToPlot ,
          Treatment == 'Safe') %>%
  mutate(Treatment = recode(Treatment, 
                            Para = 'ii) With Parasitoids',
                            Safe = 'i) Without Parasitoids') )%>%
  ggplot(aes(NicheDiff,FitnessDiff ))+
  geom_vline(xintercept = 0, linetype = 'dashed')+
  geom_ribbon(data = BoundaryValues,
              aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
  geom_ribbon(data = BoundaryValues_P,
              aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
  stat_density_2d(aes(fill = ..level..), geom = "polygon", bins = 8)+
  scale_y_log10(name = expression(paste("Fitness difference: ",kappa[i]/kappa[j])) )+
  xlab(expression(paste("Niche Difference: ", 1-rho)))+
  coord_fixed(xlim =c(-0.25, 1), ylim= c(0.2, 3))+
  geom_path(data = OneLine,
            aes(x, y), col = 'red', size = 1.5)+
  geom_line(data = BoundaryValues,
            aes(x, y2), col = 'red', size = 1.5)+
  scale_fill_gradient(low = 'blue', high= 'white', guide = FALSE,  )+
  facet_wrap(~Treatment, ncol=1)+
  theme(strip.background = element_rect(fill = 'white'),
        strip.placement = 'left',
        strip.text = element_text(hjust = 0.1, vjust = 1),
        panel.spacing = unit(2, 'lines'))-> SULSIM_Safe

SUL_SIMContourplot <- plot_grid(SULSIM_Safe, SULSIM_Para, ncol = 1 )

```


## Contour Plots
```{r fig.height=10, fig.width=6, warning=FALSE}
TidyName<-  function(name){
  parts<- str_split_fixed(name, '_', 2)  
  nicename<- paste0('italic(i):~', parts[,1], '~italic(j):~',  parts[,2])
  return(nicename)
}

sorter<-   mutate(OrderedCombs, Combination=TidyName(Combination) )
sorter$numerals <- str_replace_all(formatOL(sorter$Combination, type = 'roman'), pattern = ' ', replacement = '~')
```

```{R}
Both_TreatmentsValues %>%
  mutate(Combination=TidyName(Combination) )%>%
  left_join(sorter, by = "Combination") %>%
  filter(NicheDiff>-1)%>%
  ggplot(aes(NicheDiff,FitnessDiff ))+
  geom_ribbon(data = BoundaryValues,
              aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
  geom_ribbon(data = BoundaryValues_P,
              aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
  stat_density_2d(aes(col =Treatment),bins = 3)+
  scale_y_log10()+
  scale_colour_manual(values = c('darkred', 'blue'), labels = c('With Parasitoid', 'No Parasitoid'), name = 'Treatment')+
  facet_wrap(~Combination, ncol = 3, labeller = label_parsed)+
  coord_cartesian(xlim = c(-0.25, 1), ylim = c(0.1, 10))+
  # geom_point(data = mutate(ungroup(PosteriorLocation),Combination=TidyName(Combination) ),
  #           aes(MedianNicheDiff,MedianFitnessDiff, col = Treatment  ), shape  =4)+         # Plotting medians as little crosses 
  theme(strip.background = element_rect(fill = 'white'),
        strip.text = element_text(hjust = 0.1, vjust = 1),
        panel.spacing = unit(2, 'lines'),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank())+
  ylab(expression(paste("Fitness difference: ",kappa[i]/kappa[j])))+
  xlab(expression(paste("Niche Difference: ", 1-rho)))+
  ggtitle('c) All Species Combinations')+
  theme(legend.position = "bottom", title = element_text(hjust = 0))-> ContourPlot

ContourPlot

```

```{r fig.height=10, fig.width=10}
### Combining plots to make figure 3

LeftOfPlot<- plot_grid(Fig2_Explainer,
                       SUL_SIMContourplot,
                       ncol =1, rel_heights = c(1,2),
                       labels= c('a) Guide to coexistence plane', 'b) Example Pair: D. sulfurigaster and D. simulans'),
                       scale = 0.8)

CombinedFig2 <- plot_grid(LeftOfPlot, ContourPlot, rel_widths = c(2,3) )   
ggsave('Plots/CombinedFigure3.png', width = 10, height = 10)
ggsave('Plots/CombinedFigure3.pdf', width = 10, height = 10)

```


## All combinations as density plots

```{r fig.height=10, fig.width=7, warning=FALSE}

PlotSetOfContours<- function(Both_TreatmentsValues){
  
  Both_TreatmentsValues %>%
    mutate(Treatment = recode(Treatment, 
                              Para = 'ii) With Parasitoids',
                              Safe = 'i) Without Parasitoids') )%>%
    filter(NicheDiff>-0.5)%>%
    ggplot(aes(NicheDiff,FitnessDiff ))+
    geom_ribbon(data = BoundaryValues,
                aes(x = x, ymin = y2, ymax = y1), inherit.aes = FALSE, fill = 'grey75')+
    geom_ribbon(data = BoundaryValues_P,
                aes(x = x, ymin = y1, ymax = y2), inherit.aes = FALSE, fill = 'grey90')+
    stat_density_2d(aes(fill = ..level..), geom = "polygon")+
    scale_y_log10(limits = c(0.1, 10))+
    geom_line(data = BoundaryValues,
              aes(x, y1), col = 'red')+
    geom_line(data = BoundaryValues,
              aes(x, y2), col = 'red')+
    scale_fill_viridis_c(guide = FALSE )+
    facet_grid(CombinationOrdered~Treatment)+
    theme(strip.background = element_rect(fill = 'white'),
          strip.placement = 'left',
          strip.text = element_text(hjust = 0.1, vjust = 1),
          panel.spacing = unit(2, 'lines'))-> ContourSet
  return(ContourSet)
}

### Splitting into three plots for clarity. 
CombinationVector<-unique(Both_TreatmentsValues$Combination)

ContourSet1 <- PlotSetOfContours(filter(Both_TreatmentsValues,Combination %in% CombinationVector[1:5]))
ContourSet2 <- PlotSetOfContours(filter(Both_TreatmentsValues,Combination %in% CombinationVector[6:10]))
ContourSet3 <- PlotSetOfContours(filter(Both_TreatmentsValues,Combination %in% CombinationVector[11:15]))

ContourSet1
ContourSet2
ContourSet3

# ggsave( 'Plots/ContourSet1.pdf',ContourSet1, width = 8, height =10, units = 'in', dpi = 500 )
# ggsave( 'Plots/ContourSet2.pdf',ContourSet2, width = 8, height =10, units = 'in', dpi = 500 )
# ggsave( 'Plots/ContourSet3.pdf',ContourSet3, width = 8, height =10, units = 'in', dpi = 500 )
# 
# ggsave( 'Plots/ContourSet1.png',ContourSet1, width = 8, height =10, units = 'in', dpi = 500 )
# ggsave( 'Plots/ContourSet2.png',ContourSet2, width = 8, height =10, units = 'in', dpi = 500 )
# ggsave( 'Plots/ContourSet3.png',ContourSet3, width = 8, height =10, units = 'in', dpi = 500 )
```



