---
title: "SI 2 - Model Design - Allee Effect, Sex Ratio & Emergence Time"
author: "Chris Terry, Jinlin Chen & Owen Lewis"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(cowplot)
library(ggpubr)

# Loading raw data:
Clean_IntraData<- read_csv('Data/Clean_IntraData.csv')
CleanInterData<- read_csv('Data/CleanInterData.csv')

species <- c('BIR','PAL', 'PAN', 'PSA', 'SIM', 'SUL')

cols = c('BIR' = 'skyblue', 
         'PAL' = 'darkblue',
         'PAN' = 'red',
         'PSA'= 'darkgreen',  
         'SIM' = 'purple', 
         'SUL' = 'gold'  )
```

# Sex Ratio

```{r, message=FALSE, warning = FALSE}

Clean_IntraData %>%
  filter(Treatment  %in% c('Hot Para', 'Hot Safe'))%>%
  rowwise()%>%
  mutate(TotMales = sum(Male1, Male2, Male3, Male4, na.rm = TRUE),
         TotFemales = sum(Female1, Female2, Female3, Female4, na.rm = TRUE),
         TotalFounders= Sp1_Dens)%>%
  select(Sp = Sp1, Sp_Dens= Sp1_Dens,TotalFounders,tubeID, Treatment, TotMales, TotFemales) -> intra_sexratio

CleanInterData %>%
  filter(Treatment  %in% c('Hot Para', 'Hot Safe'),
         !is.na(FemaleSp1))%>% 
  mutate(TotMales = MaleSp1,
         TotFemales = FemaleSp1,
        TotalFounders = Sp1Dens +Sp2Dens)%>%
  select(Sp = Sp1, Sp_Dens= Sp1Dens, TotalFounders, tubeID, Treatment, TotMales, TotFemales) -> inter1_sexratio

CleanInterData %>%
  filter(Treatment  %in% c('Hot Para', 'Hot Safe'),
         !is.na(FemaleSp1))%>% 
  mutate(TotMales = MaleSp2,
         TotFemales = FemaleSp2,
        TotalFounders = Sp1Dens +Sp2Dens)%>%
  select(Sp=  Sp2, Sp_Dens= Sp2Dens, TotalFounders, tubeID, Treatment, TotMales, TotFemales) -> inter2_sexratio

AllSexRatioData <- bind_rows(intra_sexratio,inter1_sexratio, inter2_sexratio )

```

```{r, warning=FALSE, message = FALSE}
AllSexRatioData%>%
  ggplot(aes(TotMales, TotFemales))+
  geom_abline(slope=1, col = 'red', size =1.5, linetype = 'dotted')+
  geom_point(alpha = 0.5)+
  facet_wrap(~Sp)+
  coord_fixed()+ 
  scale_x_log10()+
  scale_y_log10()+
  theme(panel.background =  element_blank())+
  xlab('Males')+
  ylab('Females')
```

**Figure S2.1.** Consistent sex-ratio of emergent adults, facetted by species. Across all six species, a binomial GLM estimated only a very slight bias ($\beta$ coefficient = 0.065) towards females.   

```{r include=FALSE}
glm(  cbind(TotMales, TotFemales)~Sp* TotalFounders ,
         family = 'binomial' ,data=AllSexRatioData) %>% summary

glm(  cbind(TotMales, TotFemales)~1,
        family = 'binomial' ,data=AllSexRatioData) %>% summary
```

# Allee Effects


```{r}
Clean_IntraData%>%
  filter(Sp1_Dens <4)%>%
  filter(Treatment  %in% c('Hot Para', 'Hot Safe'))%>%
  select(Sp1, Sp1_Dens, GrandTotal, Treatment)%>%
  group_by(Sp1,Sp1_Dens, Treatment)%>%
  summarise(MeanTotalPerFounder = mean(GrandTotal/Sp1_Dens))%>%
  mutate(Treatment = recode(Treatment, 
                            'Hot Safe'='No-Parasitoid',
                            'Hot Para'= 'Parasitoid'))-> AlleeEffectData

AlleeEffectData %>%
  group_by(Sp1, Treatment)%>%
  spread(Sp1_Dens, MeanTotalPerFounder)%>%
  rename( `Per-Founder Total (Single Founder)` = `1`,
          `Per-Founder Total (Three Founders)`=   `3`) -> AlleeEffectData2

# kable(AlleeEffectData2, digits = 2)

# AlleeEffectData2 %>%
#   group_by(Treatment)%>%
#   summarise( `Mean Percentage Change For Singleton` = 100*mean((`Per-Founder Total (Single Founder)`- `Per-Founder Total (Three Founders)`)/ `Per-Founder Total (Three Founders)`)) %>%
#   kable(digits = 2)

AlleeEffectData%>%
  ggplot(aes(fill=factor(Sp1_Dens), Sp1, MeanTotalPerFounder))+
  geom_bar(stat = 'identity', position = 'dodge')+
  scale_fill_viridis_d(name = 'Number of\nFounder Pairs')+
  ylab('Mean Total Offspring\nPer Founder Pair')+
  xlab('Species')+
  facet_wrap(~Treatment)
  
```

**Figure S2.2.** Per-female offspring is consistently lower for singleton founder pairs than triplets in the monoculture trials.

# Emergence Time Model

Model used to infer mean emergence time from starting densities, based on counts from the single-species tubes, was selected based on backwards stepwise-selection with AIC. 

```{r}
Clean_IntraData%>%
  arrange(tubeID)%>%
  filter(Temp == 'Hot')%>%
  mutate(Dens0 = Sp1_Dens,
         SpID = Sp1)%>%
  filter(GrandTotal >0) -> EmergenceTimeData

lm(data = EmergenceTimeData,
   MeanEmergence ~ Dens0 * SpID*Treatment)-> Mod_StartCount_all

tmp<-step(Mod_StartCount_all) #  Suggests dropping tree-way interaction

Mod_StartCount <- lm(data = EmergenceTimeData,
   MeanEmergence ~ Dens0 + SpID + Treatment + Dens0:SpID + Dens0:Treatment + SpID:Treatment)

summary(Mod_StartCount)

lm(data = EmergenceTimeData,
   MeanEmergence ~ GrandTotal * SpID*Treatment)-> Mod_EndCount
# Using end counts is similar accuracy, but harder to extrapolate

#summary(Mod_EndCount)
```

```{r}
Predictions<- predict(Mod_StartCount,EmergenceTimeData )

SingleSpeciesPredictionData <-data.frame('Predicted Emergence' = Predictions,
           'Observed Mean Emergence' = EmergenceTimeData$MeanEmergence,
           'Sp' = EmergenceTimeData$Sp1,
           'Density' = EmergenceTimeData$Dens0,
           'Treatment' =EmergenceTimeData$Treatment )%>%
  mutate(Treatment = recode(Treatment, 
                             'Hot Safe'='No-Parasitoid',
                            'Hot Para'= 'Parasitoid'))

SingleSpeciesPredictionData %>%
  ggplot(aes(Predicted.Emergence,Observed.Mean.Emergence, col=Sp ))+
  geom_point()+
  geom_abline(slope=1)+
  facet_wrap(~Treatment)+
  coord_fixed()+
  scale_colour_manual(values = cols, name = 'Species')+
  ylab('Observed Mean Emegence (days)')+
  xlab('Predicted Emergence (days)')

```

**Figure S2.3** Quality of fit of emergence time model for monocultures, facetted by treatment. 

```{r}


SingleSpeciesPredictionData%>%
  ggplot(aes(y= Observed.Mean.Emergence, x=Density,  col=Sp ))+
  geom_point()+
  geom_line(aes(y= Predicted.Emergence))+
  facet_wrap(~Treatment)+
  ggtitle('Emergence Time vs Starting Density')+
  scale_colour_manual(values = cols, name = 'Species')+
 ylab('Observed Mean Emegence (days)')+
  xlab('Starting Density')


```

**Figure S2.4** Effect of number of founders on emergence time. 



